---
title: 'Lab 9, Part 1: Time Series Analysis'
author: "Sydney Mayes"
date: "2023-03-10"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
library(tidyverse)
library(here)
library(tsibble)
library(feasts)
library(fable)
library(broom)
```

## Read in our energy data and convert to a tsibble
Data here are from teh US Energy Infromation Administration. US Residential energy consumption from January 1973 through October 2017. Units are in trillions of btus

```{r}
energy_df <- read_csv(here('data', 'energy.csv'))

energy_ts <- energy_df %>% 
  mutate(date = tsibble::yearmonth(month)) %>% 
  as_tsibble(key = NULL, index = date)

```

## Exploratory time series viusualization

```{r}
ggplot(data = energy_ts, aes(x = date, y = res_total)) +
  geom_line() +
  labs(y = 'Residential energy consumption\n (Trillion BTU)')

# There seems to be an increasing trend, but may level off around 2005. There is a pretty clear seasonality -- a big spike that seems to match up well with each year. A big spike then a small spike rather than a valley. Doesn't seem to be strong cyclic nature or outliers.
```
### Season plot

```{r}
energy_ts %>% 
  gg_season(y = res_total) + # from feasts package; already knows the x
  theme_minimal() +
  scale_color_viridis_c() + # c is for continuous variable since dealing with years
  labs(x = 'month', 
       y = 'residential energy consumption (trillion BTU)')

# Based on the graph it looks like November-February higher consumption may be related to winter months/less daylight/heat. Summer increase may be due to AC usage.
```

### Subseries plot
```{r}
energy_ts %>% 
  gg_subseries(y = res_total) 

# shows each month and the trend across the whole time series
```
## Decomposition (into different components)

```{r}
# Classical decomposition based on trend; seasonality is average across all years
# But what we used last week is STL decomposition. Loess is the method for estimating nonlinear relationships, based on weighted rolling average; very different from classical decomposition

# STL decomposition vs classical decomposition -- classical has same seasonality over entier series, but STL allows seasonality to change over time. 

### Calculate the STL decomposition:
dcmp <- energy_ts %>% 
  model(STL(res_total ~ season()))

### Visualize
components(dcmp) %>% 
  autoplot() + # autoplot adds a ggplot object so use + not pipe operator
  theme_minimal()
# The vertical bars on the plot show the height of the remainder portion (in this case about 600 units for each of the 4 graphs)
  
```
## Autocorrelation function (ACF)
```{r}
# Compares the value in every month to the value in one month later, two months later, three, etc

energy_ts %>% 
  ACF(y = res_total) %>%  # feasts function over stats since we have a tsibble (time series)
  autoplot()

# 6 months out has a small positive correlation probably because of the summer AC usage. We see similar trends every 12 months. This is in line with what we expected from previous plots
```
### Forecasting with Holt-Winters exponential smoothing

```{r}
# pays more attention to recent observations/events than those in the far past. Exponential decay in terms of the strength of the contributions of past events

### Create our model
energy_fit <- energy_ts %>% 
  model(ets = ETS(res_total ~ season('M'))) # will consider this as a multiplicative seasonality model

### create a forecast based on our model, 10 years into the future
energy_forecast <- energy_fit %>% 
  forecast(h = '10 years') # from the fable package

energy_forecast %>% 
  autoplot()

# confidence intervals get wider out in time as the years go out since there is more uncertainty
```
Now plot the actual res_total values on top of the predicted values to see how well our fit data matches up with our observed data
```{r}
### append the predicted values (and residuals) to the original energy data 
energy_predicted <- broom::augment(energy_fit)

ggplot(data = energy_predicted) + 
  geom_line(aes(x = date, y = res_total), color = 'black') +
  geom_line(aes(x = date, y = .fitted), color = 'red', alpha = 0.7)

# red lines are predicted, black is observed. Looks like a fairly good fit. 

# now let's check out the residuals to make sure they are centered around zero, normally distributed around 0, mean of zero
ggplot(data = energy_predicted, aes(x = .resid)) +
  geom_histogram()
# yes they are

# Are there any correlations or patterns of residuals? If so there may be info we didnt account for when predicting
ggplot(data = energy_predicted, aes(x = date, y = .resid)) +
  geom_point()
# looks to be pretty uncorrelated from the plot, yay

# The above are all important things to check on the model. 

# In the key there is extra code about forecasting methods including seasonal naive, etc. 
```

